---
title: "Heston Calibration"
author: "Matthew Dixon"
date: "February 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r}
require("quantmod")
require("xts")
source('HestonFourierCosine.R')
source('getChain.R')
```

## Problem statement

Begin by loading the data

```{r echo=FALSE}
ZNGA <- getChain('data/ZNGA.csv')
ZNGA$timestamp<- as.POSIXct(as.numeric(as.character(ZNGA$timestamp)), origin='1970-01-01', tz='GMT')
ATM_CALL_PRICE <- as.xts(ZNGA$mid_price, order.by = ZNGA$timestamp,frequency = NULL)
UNDERLYING <- as.xts(ZNGA$underlying, order.by = ZNGA$timestamp,frequency = NULL) 
chartSeries(ATM_CALL_PRICE, TA=list("addTA(UNDERLYING, col=4)"))
```

Perform the calibration. Let's use the first 50 data points for the calibration and lets see how good the Heston model fit over the remaining 50 data points.

Questions:
(1) how stable is the calibration?
(2) how sensitive is the calibration to the choice of initial parameter values?
(3) what is the better of calibrating more frequently - does this lead to a better model fit?

```{r echo=FALSE}
source('logdensity2loglik.R')
source('logdensity2info.R')
source('HestonFourierCosine.R')
source('getImpliedVolatility.R')
source('mymle.R')
source('models/ModelB4.R')
require(nloptr)
require(pracma)

args<-list(maxiter=50, eps=1e-8, print_level=3)
eps <-1e-8
#rho, kappa, theta, sigma
args$l <- c(-1.0 + eps,eps,eps, eps)
args$u <- c(1.0-eps,4.0-eps,1.0-eps,2.0-eps)
#rho, kappa, theta, sigma
eval_g_ineq <- function (x) {
  grad <- c(0,-2.0*x[3],-2.0*x[2],2.0*x[4])
  return(list("constraints"=c(x[4]*x[4] - 2.0*x[2]*x[3]), "jacobian"=grad))  
}
# Heston model:
# dln(S_t) = \mu dt + \sqrt{V_t}dW_t^{1}
# dV_t = \kappa(\theta - V_t)dt  + \sigma \sqrt{V_t}dW_t^{2}
# 4 parameters to be estimated: (rho, kappa, theta, sigma)  
  

S_0      <- ZNGA$UNDERLYING[1] 
v_0      <- 0.4  # the initial variance is assumed to be unknown
rate     <- 0.1  # the risk free rate is assumed to be known
q        <- 0

a_0      <- rate -q
b_0      <- -0.5
rho_0    <- -0.8
kappa_0  <-3
theta_0  <-0.2
sigma_0  <-0.25 

param_0<-c(rho_0,kappa_0,theta_0,sigma_0)

args$mode = 'option' # calibration to ATM option prices 

# daily data: del = 1/52
# 30 second updates
delta <- 0.5*(1/252)/(60*24)
n <- dim(ZNGA)[1]
n_train <- 50
# s_t=ln(S_t)
T_0 <-ZNGA$maturity[1]
K_0 <-ZNGA$strike[1]
args$callput = ZNGA$type[1]
x1 <- log(ZNGA$underlying[1:n_train])
x2 <- rep(v_0, n_train)
V <-  ZNGA$mid_price[1:n_train] 
x <- cbind(x1,x2,V)
```
Now calibrate the model
```{r echo=FALSE}
output <- mymle(ModelHeston, x, delta, param_0, args)
print(output$se)
print (output$se_robust)
rho   <- output$solution[1]
kappa <- output$solution[2]
theta <- output$solution[3]
sigma <- output$solution[4]
v     <- getImpliedVolatilty(ZNGA$mid_price[n_train],v_0, args)
```
Now evaluate the Heston model over the remaining period
```{r echo=FALSE}
ZNGA$model_price <- rep(0,n)
ZNGA$vega <- rep(0,n)
for (i in (n_train+1):n){
  ZNGA$model_price[i] <- HestonCOS(ZNGA$underlying[i],K_0,T_0,rate,q,sigma,kappa,theta,v,rho,args$callput)
  ZNGA$vega[i] <- HestonCOS_vega(ZNGA$underlying[i],K_0,T_0,rate,q,sigma,kappa,theta,v,rho,args$callput)
}

```

```{r echo=FALSE}
#plot the results
posix_timestamp<- as.POSIXct(as.numeric(ZNGA$timestamp[(n_train+1):n]), origin='1970-01-01', tz='GMT')
mid<-as.xts(ZNGA$mid_price[(n_train+1):n],order.by= posix_timestamp,frequency = NULL) 

model<-as.xts(ZNGA$model_price[(n_train+1):n], order.by=posix_timestamp,frequency = NULL)
chartSeries(mid, TA="addTA(model, col=4)")

```